---
title: "ESM 244 Lab"
author: "An Bui"
date: "2/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
# General packages
library(tidyverse)
library(janitor)
library(plotly)
library(RColorBrewer)

# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)

# Packages for text mining/sentiment analysis/word cloud
library(pdftools)
library(tidytext)
library(wordcloud)
```

#### Part 1. k-means clustering using `iris` dataset
Recall: uses a number of randomly placed centroid locations and iteratively converges to the centroids
```{r}
#### use janitor to clean up column names ####
iris_nice <- iris %>% 
  clean_names()

#### quick plot ####
ggplot(iris_nice) +
  geom_point(aes(x = petal_length, 
                 y = petal_width, 
                 color = species))
# looks like 3 clusters because of colors, but if you didn't have those, it'd look like 2
```

How many clusters should exist (according to R)? Use `NbClust` function to determine the best number of clusters.

```{r determine number of clusters}
                      # columns of data to be clustered
number_est <- NbClust(iris_nice[1:4], 
                      # minimum number of clusters
                      min.nc = 2, 
                      # maximum number of clusters
                      max.nc = 10, 
                      # clustering method
                      method = "kmeans")

# * According to the majority rule, the best number of clusters is  2 
# but you know that there are 3 species, so 2 might not make sense so stick with 3 clusters when performing kmeans
```

Perform k-means clustering with 3 groups.

```{r}
iris_km <- kmeans(iris_nice[1:4], 
                  # how many clusters to exist in final grouping
                  3)

# look at how many observations exist in each cluster
iris_km$size

# multivariate center location associated with these three clusters
iris_km$centers

# what cluster has each observation been assigned to?
iris_km$cluster
# clusters are fairly well separated; clear groupings without much overlap

#### bind cluster assignment to original data ####
iris_cl <- data.frame(iris_nice, 
                      # new column cluster_no which has what cluster each observation has been assigned to                           (set as a factor)
                      cluster_no = factor(iris_km$cluster))

#### basic ggplot ####
ggplot(iris_cl) +
  geom_point(aes(x = sepal_length, y = sepal_width, color = cluster_no))
# there are observations that overlap between different clusters because clusters are assigned with multivariate data, but only representing 2 variables in the plot

#### better ggplot ####
ggplot(iris_cl) +
  geom_point(aes(x = petal_length,
                 y = petal_width,
                 color = cluster_no,
                 pch = species)) +
  scale_color_brewer(palette = "Set2")

#### use plotly to make a 3D representation of clusters ####
plot_ly(x = iris_cl$petal_length,
        y = iris_cl$petal_width,
        z = iris_cl$sepal_width,
        # plot type
        type = "scatter3d",
        # color points by cluster_no
        color = iris_cl$cluster_no,
        # symbol by species
        symbol = iris_cl$species,
        colors = "Set1")
```

#### Part 2. Hierarchical cluster analysis

```{r}
#### data ####
wb_env <- read_csv("wb_env.csv")

#### wrangling ####
wb_ghg_20 <- wb_env %>% 
  arrange(-ghg) %>% 
  head(20)

#### scale variables, then coerce back to dataframe ####
# scale is in baseR
# select columns to scale
wb_scaled <- as.data.frame(scale(wb_ghg_20[3:7]))
# data frame is still associated with rows that were previously there

# add rownames
rownames(wb_scaled) <- wb_ghg_20$name

#### create dissimilarity matrix using dist() in baseR ####
diss <- dist(wb_scaled, method = "euclidian")
# output is in format dist

#### hierarchical agglomerative clustering by complete linkage ####
hc_complete <- hclust(diss, method = "complete")
plot(hc_complete)

#### divisive analysis clustering ####
# everything is its own group to start with
hc_divisive <- diana(diss)
plot(hc_divisive)

#### dendrogram ####
dend1 <- as.dendrogram(hc_complete)
dend2 <- as.dendrogram(hc_divisive)

tanglegram(dend1, dend2)

#### use ggdendro package ####
ggdendrogram(hc_complete,
             rotate = TRUE) +
  theme_minimal()
```

#### Part 3. Intro to text analysis

```{r set up}
# specify a file path
greta_thunberg <- file.path("greta_thunberg.pdf")

# read in text
thunberg_text <- pdf_text(greta_thunberg)
```

```{r wrangling}
# entire thing exists in one line
thunberg_df <- data.frame(text = thunberg_text) %>% 
  # split up text by \\n
  mutate(text_full = str_split(text, '\\n')) %>% 
  unnest(text_full)

speech_text <- thunberg_df %>% 
  select(text_full) %>% 
  slice(4:18)

sep_words <- speech_text %>% 
  unnest_tokens(word, text_full)
# every word now exists in its own row

word_count <- sep_words %>% 
  count(word, sort = TRUE) 

words_stop <- sep_words %>% 
  anti_join(stop_words)
# filler words taken out
```

```{r analysis}
#### sentiment word list ####
pos_words <- get_sentiments("afinn") %>% 
  filter(score == 5 | score == 4) %>% 
  head(20)

neutral_words <- get_sentiments("afinn") %>% 
  filter(score, -1, 1) %>% 
  head(20)

neg_words <- get_sentiments("afinn") %>% 
  filter(score == -5 | score == -4) %>% 
  head(20)

#### bind ####
sent_afinn <- words_stop %>% 
  # inner join only keeps things with matches in both
  inner_join(get_sentiments("afinn"))

#### nrc sentiments ####
sent_nrc <- words_stop %>% 
  inner_join(get_sentiments("nrc"))

nrc_count <- sent_nrc %>% 
  group_by(sentiment) %>% 
  tally()

#### create a word cloud ####
wordcloud(word_count$word,
          freq = word_count$n,
          min.freq = 1,
          max.words = 65,
          scale = c(2, 0.1),
          colors = brewer.pal(3, "Dark2"))
```













